Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
Complete negative example for checking and amend consistency function

We will consider both a consistent and an inconsistent example to show how
checking works. The following is an example with a looser but still consistent
version of the \textit{fivepoint} stencil specification inferred previously in
the Subsection~\ref{subsec:inference}:
%
\begin{minted}[breakindent=8.1em]{fortran}
!= region fivepoint = centered(depth=1,dim=1) + centered(depth=1,dim=2)
!= stencil readOnce, fivepoint :: b
a(i,j) = b(i, j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j)
\end{minted}
%
\begin{enumerate}
\item Generate model from the specification as described in
  Section~\ref{sec:semantics}.
%
\begin{dmath*}
\interp{\texttt{stencil readOnce, fivepoint}}_2 =
  \textsf{once}(
    \textsf{exact}(
      \{ \vtwo{-1}{\infty},
          \vtwo{0}{\infty},
          \vtwo{1}{\infty},
          \vtwo{\infty}{-1},
          \vtwo{\infty}{0},
          \vtwo{\infty}{1}\}))
\end{dmath*}
%
\item Now $cons$ function is applied with this model along side with the
  assignment on line three and array $b$. For brevity the assignment is
  referred as $L3$ and the set of unifiers defined in the previous step as
  $m$.

@@

  The unifications for the lower bound:
  \begin{align*}
     \vtwoh{0}{-1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{-1}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{\infty}{0}
  \end{align*}

  The unifications for the upper bound:
  \begin{align*}
     \vtwoh{-1}{0} & \sim_2 \vtwoh{-1}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{1}{\infty}  \\
     \vtwoh{0}{-1} & \sim_2 \vtwoh{\infty}{-1} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{\infty}{1}
  \end{align*}

@@

>>>>

We will consider an extended example of a stencil computation that is
consistent with its annotation to show how checking works. The following is an
example with a looser but still consistent version of the \textit{fivepoint}
stencil specification inferred previously in the
Subsection~\ref{subsec:inference}:
%
\begin{minted}[breakindent=8.1em]{fortran}
!= region fivepoint = centered(depth=1,dim=1) + centered(depth=1,dim=2)
!= stencil readOnce, fivepoint :: b
a(i,j) = b(i, j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j)
\end{minted}
%
\begin{enumerate}
\item Generate model from the specification as described in
  Section~\ref{sec:semantics}.
%
\begin{dmath*}
\interp{\texttt{stencil readOnce, fivepoint}}_2 =
  \textsf{once}(
    \textsf{exact}(
      \{ \vtwo{-1}{\infty},
          \vtwo{0}{\infty},
          \vtwo{1}{\infty},
          \vtwo{\infty}{-1},
          \vtwo{\infty}{0},
          \vtwo{\infty}{1}\}))
\end{dmath*}
%
\item Now $cons$ function is applied with this model along side with the
  assignment on line three and array $b$. For brevity the assignment is
  referred as $L3$ and the set of unifiers defined in the previous step as
  $m$.

@@

  The unifications for the upper bound:
  \begin{align*}
     \vtwoh{0}{-1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{-1}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{\infty}{0}
  \end{align*}

  The unifications for the lower bound:
  \begin{align*}
     \vtwoh{-1}{0} & \sim_2 \vtwoh{-1}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{1}{\infty}  \\
     \vtwoh{0}{-1} & \sim_2 \vtwoh{\infty}{-1} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{\infty}{1}
  \end{align*}

@@

Now when $L3$ is slightly changed with an additional term \texttt{b(i+2,j)}, the
new assignment $L3'$ becomes:
%
\begin{minted}[firstnumber=3]{fortran}
a(i,j) = b(i, j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j) + b(i+2,j+2)
\end{minted}
%
The first four steps will proceed the same way and in the fifth step result
of $\mathit{analyse}(L3',b)$ will have the extra term $\vtwoh{2}{2}$. In sixth
step, since the model has not changed and the relativised indices are only
augmented with a new element, the unification for lower bound still hold.
However, there are not elements in the model, $m$, that can unify with
$\vtwoh{2}{2}$, hence the model fails to hold as an upper bound.

Note that if the specification was as follows:
%
\begin{minted}{fortran}
!= stencil atLeast, readOnce, fivepoint :: b}
\end{minted}
%
The consistency function would have held as
$\mathit{cons}(\textsf{mult}(\textsf{up}(m)),L3',b)$ for the upper bound would
have never been needed. This semantically makes sense since the region
$\texttt{fivepoint}$ is indeed a lower bound for the access pattern of $b$.
